Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?

### 2020-12-26 docs/00-manuscripts/ieee/ieee-manuscript.tex
add more information how the data surveyed
Explain why NDVI was chosen as the base for NRI calculation

The \textit{Pinus radiata} plots of this study, namely Laukiz1, Laukiz2, Luiando and Oiartzun, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
Oiartzun has the most observations (n = 559) while Laukiz2 shows the largest area size (1.44 ha).
All plots besides Luiando are located within 100 km from the coast (\autoref{fig:study_area}).
In total 1808 observations are available (Laukiz1 = 559, Laukiz2 = 451, Luiando = 301, Oiartzun = 497).
Field surveys were conducted in September 2016.

@@

To use the full potential of the hyperspectral data, all possible vegetation indices supported by the R package hsdar (89 in total) as well as all possible \ac{NRI} combinations were calculated from the reflectances.
The following formula was used for the NRI calculation:

@@

@@

The best aggregated performance of this study (SVM + \enquote{Info Gain} filter, RMSE 27.99 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be around 19 percentage points.
However, leaving out a single plot would also change the prediction results for the other plots because the observations from Laukiz2 would not be available for model training.
Due to the apparent presence of model overfitting in this study it can be postulated that more training data representing a greater variety of situations is needed.
A model can only make robust predictions if it has learned relationships across the whole range of the response.
Hence, care should be taken when predicting to the landscape scale using models fitted on this dataset due to their lack of generalizability caused by the limitations of the available training data.
However, when inspecting the fold level performances, it can be concluded that the model performed reasonably well predicting defoliation greater than 50\% but failed for lower levels.
This applied to all learners of this study.

@@

% parameter limits
\begin{table}[b!]
	\centering
	\caption[t]{Hyperparameter ranges and types for each model.
		Hyperparameter notations from the respective R packages were used.}
	\begingroup\scriptsize
	\begin{tabularx}{.5\textwidth}{llXXrXX}
		\\
		\specialcell{Model                                              \\ (package)}                & Hyperparameter              & Type    & Start     & End      & Default \\
		\toprule
		\multirow{3}{*}{\specialcell{RF                                 \\ (ranger)}}       & \texttt{$x_try$}    & dbl & 0         & 0.5      & -       \\
		 & \texttt{min.node.size}      & int & 1        & 10      & 1   \\
		 & \texttt{sample.fraction}    & dbl & 0.2      & 0.9     & 1   \\
		\midrule
		\multirow{2}{*}{\specialcell{SVM                                \\ (kernlab)}}     & \texttt{C}                  & dbl & $2^{-10}$ & $2^{10}$ & 1       \\
		 & \texttt{$\sigma$}           & dbl & $2^{-5}$ & $2^{5}$ & 1   \\
		\midrule
		\multirow{7}{*}{\specialcell{XGBoost                            \\ (xgboost)}} & \texttt{nrounds}            & int & 10        & 600      & -       \\
		 & \texttt{colsample\_bytree}  & dbl & 0.3      & 0.7     & 1   \\
		 & \texttt{subsample}          & dbl & 0.25     & 1       & 1   \\
		 & \texttt{max\_depth}         & int & 1        & 10      & 6   \\
		 & \texttt{gamma}              & int & 0        & 10      & 0   \\
		 & \texttt{eta}                & dbl & 0.01     & 0.6     & 0.3 \\
		 & \texttt{min\_child\_weight} & int & 0        & 20      & 1   \\
		\bottomrule
	\end{tabularx}
	\endgroup
	\label{tab:hyperparameter_limits}
\end{table}

>>>>

The \textit{Pinus radiata} plots of this study, namely Laukiz1, Laukiz2, Luiando and Oiartzun, are located in the northern part of the Basque Country (\autoref{fig:study_area}).
Oiartzun has the most observations (n = 559) while Laukiz2 shows the largest area size (1.44 ha).
All plots besides Luiando are located within 100 km from the coast (\autoref{fig:study_area}).
In total 1808 observations are available (Laukiz1 = 559, Laukiz2 = 451, Luiando = 301, Oiartzun = 497).
Field surveys were conducted in September 2016.
Defoliation was measured via visual inspection using 5\% intervals.
No estimation error was recorded.

@@

To use the full potential of the hyperspectral data, all possible vegetation indices supported by the R package hsdar (89 in total) as well as all possible \ac{NRI} combinations were calculated from the reflectances.
The following NDVI-based formula was used for the NRI calculation:

@@

The choice of NDVI serving as the base formula for NRI calculation was determined by the implementation of the \texttt{nri} function in the \texttt{hsdar} R package.

@@

The best aggregated performance of this study (SVM + \enquote{Info Gain} filter, RMSE 27.99 p.p.) has to be seen in the light of model overfitting (see \autoref{subsec:perf-plot-char}).
Leaving out the performance on Laukiz2 when aggregating results, the mean RMSE would be around 19 percentage points.
However, leaving out a single plot would also change the prediction results for the other plots because the observations from Laukiz2 would not be available for model training.
Due to the apparent presence of model overfitting in this study it can be postulated that more training data representing a greater variety of situations is needed.
A model can only make robust predictions if it has learned relationships across the whole range of the response.
Hence, care should be taken when predicting to the landscape scale using models fitted on this dataset due to their lack of generalizability caused by the limitations of the available training data.
However, when inspecting the fold level performances, it can be concluded that the model performed reasonably well predicting defoliation greater than 50\% but failed for lower levels.
This applied to all learners of this study.
The overall performance achieved in this study can be classified as "fair".

@@

% parameter limits
\begin{table}[h!]
	\centering
	\caption[t]{Hyperparameter ranges and types for each model.
		Hyperparameter notations from the respective R packages were used.}
	\begingroup\scriptsize
	\begin{tabularx}{.5\textwidth}{llXXrXX}
		\\
		\specialcell{Model                                              \\ (package)}                & Hyperparameter              & Type    & Start     & End      & Default \\
		\toprule
		\multirow{3}{*}{\specialcell{RF                                 \\ (ranger)}}       & \texttt{$x_try$}    & dbl & 0         & 0.5      & -       \\
		 & \texttt{min.node.size}      & int & 1        & 10      & 1   \\
		 & \texttt{sample.fraction}    & dbl & 0.2      & 0.9     & 1   \\
		\midrule
		\multirow{2}{*}{\specialcell{SVM                                \\ (kernlab)}}     & \texttt{C}                  & dbl & $2^{-10}$ & $2^{10}$ & 1       \\
		 & \texttt{$\sigma$}           & dbl & $2^{-5}$ & $2^{5}$ & 1   \\
		\midrule
		\multirow{7}{*}{\specialcell{XGBoost                            \\ (xgboost)}} & \texttt{nrounds}            & int & 10        & 600      & -       \\
		 & \texttt{colsample\_bytree}  & dbl & 0.3      & 0.7     & 1   \\
		 & \texttt{subsample}          & dbl & 0.25     & 1       & 1   \\
		 & \texttt{max\_depth}         & int & 1        & 10      & 6   \\
		 & \texttt{gamma}              & int & 0        & 10      & 0   \\
		 & \texttt{eta}                & dbl & 0.01     & 0.6     & 0.3 \\
		 & \texttt{min\_child\_weight} & int & 0        & 20      & 1   \\
		\bottomrule
	\end{tabularx}
	\endgroup
	\label{tab:hyperparameter_limits}
\end{table}

\section{Vegetation index formulas}
